---
title: "Install Shiny-phyloseq"
output: html_document
---

If you are reading this, it is because you are interested
in launching/running an instance of Shiny-phyloseq.
The Shiny-phyloseq "front end", the GUI web application,
will run on any modern web browser.
Note that older versions of Microsoft's Internet Explorer are **not supported**.
Firefox, Chrome, Opera, Safari and others should be fine
for connecting to a Shiny-phyloseq session.

For running the Shiny-phyloseq "back end", you must have
the latest version of R installed on your system
as well as several additional R packages.

## Auto-Install Shiny-phyloseq (Recommended)

Simply launching Shiny-phyloseq should also install missing/old packages.

The following R code will launch Shiny-phyloseq on most systems.

```{r eval=FALSE}
install.packages("shiny")
shiny::runGitHub("shiny-phyloseq","joey711")
```


## Manual Install of Shiny-phyloseq Dependencies

The following R code will install required packages on your local system.
Note that this assumes that you have an active internet connection and updated R installation.

```{r, eval=FALSE}
source("https://raw.githubusercontent.com/joey711/shiny-phyloseq/master/install.R")
```

This line executes the same package install/update code that is used by the auto-installer in Shiny-phyloseq itself, but as a one-time batch event that you can run in advance. This is only recommended if you have never launched Shiny-phyloseq before and want to get the pakcage installation out of the way. This also provides the exact recommended dependency and installation code, in case your system has special requirements -- like special development versions of packages, or a non-standard installation of R. For most systems, this is not necessary.

